---
title: "Hatching_fractions"
author: "Lieke Vlaar"
date: "30/11/2021"
output: html_document
---

Necessary packages are loaded
* ggplot2
* patchwork
* ggpmisc
* extrafont
* tidyverse
* dplyr
* plotrix
* stringr

```{r include = FALSE}
if ("knitr" %in% installed.packages()){
  library("knitr")
} else {
  install.packages("knitr")
  library("knitr")
}
```

```{r include=FALSE}
#########
# Library
#########
if ("checkpoint" %in% installed.packages()){
  library("checkpoint") # https://cran.r-project.org/web/packages/checkpoint/index.html
} else {
  install.packages("checkpoint")
  suppressPackageStartupMessages(library("checkpoint"))
  suppressPackageStartupMessages(checkpoint("2021-11-30"))  # all packages in your project will be taken from that date.
}
if ("ggplot2" %in% installed.packages()){
  library("ggplot2")
} else {
  BiocManager::install("ggplot2")
  library("ggplot2")
}
if ("patchwork" %in% installed.packages()){
  library("patchwork")
} else {
  BiocManager::install("patchwork")
  library("patchwork")
}
if ("ggpmisc" %in% installed.packages()){
  library("ggpmisc")
} else {
  BiocManager::install("ggpmisc")
  library("ggpmisc")
}
if ("extrafont" %in% installed.packages()){
  library("extrafont")
} else {
  BiocManager::install("extrafont")
  library("extrafont")
}
if ("tidyverse" %in% installed.packages()){
  library("tidyverse")
} else {
  BiocManager::install("tidyverse")
  library("tidyverse")
}
if ("dplyr" %in% installed.packages()){
  library("dplyr")
} else {
  BiocManager::install("dplyr")
  library("dplyr")
}
if ("plotrix" %in% installed.packages()){
  library("plotrix")
} else {
  BiocManager::install("plotrix")
  library("plotrix")
}
if ("stringr" %in% installed.packages()){
  library("stringr")
} else {
  BiocManager::install("stringr")
  library("stringr")
}
```

# Get hatching per fraction

Load required functions

```{r loadfunctions include = F}
source("scripts/hatch_raw.R")
source("scripts/hatch_calc.R")
source("scripts/hatch_summ.R")
```

Plot the hatching per fraction per genotype:

```{r plothatchingperfr, fig.height = 8, fig.width = 5}
hatch_r <- hatch_raw(input = "input/210810_frac.csv")
hatch_c <- hatch_calc(input = "input/210810_frac.csv")
hatch_s <- hatch_summ(input = "input/210810_frac.csv")
hatch_s_A <- hatch_s[grep("A", hatch_s$treatment), ]
hatch_s_D <- hatch_s[grep("D", hatch_s$treatment), ]
hatch_s_S <- hatch_s[grep("S", hatch_s$treatment), ]
hatch_c_A <- hatch_c[grep("A", hatch_c$treatment), ]
hatch_c_D <- hatch_c[grep("D", hatch_c$treatment), ]
hatch_c_S <- hatch_c[grep("S", hatch_c$treatment), ]
hatch_s_W <- hatch_s[grep("water", hatch_s$treatment),]
hatch_c_W <- hatch_c[grep("water", hatch_c$treatment),]

hatch_s_A <- rbind(hatch_s_A, hatch_s_W)
hatch_c_A <- rbind(hatch_c_A, hatch_c_W)
hatch_s_D <- rbind(hatch_s_D, hatch_s_W)
hatch_c_D <- rbind(hatch_c_D, hatch_c_W)
hatch_s_S <- rbind(hatch_s_S, hatch_s_W)
hatch_c_S <- rbind(hatch_c_S, hatch_c_W)

g_A <- ggplot(hatch_s_A, aes(x = treatment, y = mean_hatch)) +
  geom_bar(stat = "identity", fill = "white", colour = "black") +
  theme_minimal() +
  geom_errorbar(aes(ymin=mean_hatch-sterr, ymax=mean_hatch+sterr), width=.2) + 
  geom_jitter(data = hatch_c_A, aes(x = treatment, y = hatchingpercent), width = 0.1) +
  scale_x_discrete(name = "") +
  scale_y_continuous(name = "% hatch", limits = c(-10,74), breaks = c(0,20,40,60,80)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1.5, hjust=1.25))
g_D <- ggplot(hatch_s_D, aes(x = treatment, y = mean_hatch)) +
  geom_bar(stat = "identity", fill = "white", colour = "black") +
  theme_minimal() +
  geom_errorbar(aes(ymin=mean_hatch-sterr, ymax=mean_hatch+sterr), width=.2) + 
  geom_jitter(data = hatch_c_D, aes(x = treatment, y = hatchingpercent), width = 0.1) +
  scale_x_discrete(name = "") +
  scale_y_continuous(name = "% hatch", limits = c(-10,74), breaks = c(0,20,40,60,80)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1.5, hjust=1.25))
g_S <- ggplot(hatch_s_S, aes(x = treatment, y = mean_hatch)) +
  geom_bar(stat = "identity", fill = "white", colour = "black") +
  theme_minimal() +
  geom_errorbar(aes(ymin=mean_hatch-sterr, ymax=mean_hatch+sterr), width=.2) + 
  geom_jitter(data = hatch_c_S, aes(x = treatment, y = hatchingpercent), width = 0.1) +
  scale_x_discrete(name = "") +
  scale_y_continuous(name = "% hatch", limits = c(-10,74), breaks = c(0,20,40,60,80)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1.5, hjust=1.25))

g_A / g_D / g_S
```

# Metabolomics of fractions

Import, filter and trim the dfs (pos and neg)

```{r}
fr_pos <- read.csv("input/fractions_pos.csv", header = TRUE, fileEncoding = "UTF-8-BOM")
fr_neg <- read.csv("input/fractions_neg.csv", header = TRUE, fileEncoding = "UTF-8-BOM")
compounds_fr_pos <- fr_pos[,1:13]
compounds_fr_neg <- fr_neg[,1:13]
fr_pos_b <- fr_pos[,colnames(fr_pos) == "blank"]
fr_neg_b <- fr_neg[,colnames(fr_neg) == "blank"]
fr_pos <- fr_pos[,-c(1:14)]
fr_neg <- fr_neg[,-c(1:14)]
fr_pos_f <- fr_pos[(!apply(fr_pos<(fr_pos_b*1.1),1,all)),]
fr_neg_f <- fr_neg[(!apply(fr_neg<(fr_neg_b*1.1),1,all)),]
```

Do transformations (centre and scale) for PCA plot. There is no vsn transformation since there too little sample to base this transformation on.Then plot pca plots.
```{r}
fr_pos_ft <- prep.autoscale(t(fr_pos_f), center = TRUE, scale = TRUE)
fr_neg_ft <- prep.autoscale(t(fr_neg_f), center = TRUE, scale = TRUE)
meanSdPlot(t(fr_pos_f))
meanSdPlot(fr_pos_ft)
meanSdPlot(t(fr_neg_f))
meanSdPlot(fr_neg_ft)

source("scripts/plotpca.R")
plotpca2(x = fr_neg_ft) + scale_size(guide = "none")
plotpca2(x = fr_pos_ft) + scale_size(guide = "none")
```
