---
title: "Hatching_fractions"
author: "Lieke Vlaar"
date: "30/11/2021"
output: html_document
---

Necessary packages are loaded
* ggplot2
* patchwork
* ggpmisc
* extrafont
* tidyverse
* dplyr
* plotrix
* stringr

```{r include = FALSE}
if ("knitr" %in% installed.packages()){
  library("knitr")
} else {
  install.packages("knitr")
  library("knitr")
}
```

```{r include=FALSE}
#########
# Library
#########
if ("checkpoint" %in% installed.packages()){
  library("checkpoint") # https://cran.r-project.org/web/packages/checkpoint/index.html
} else {
  install.packages("checkpoint")
  suppressPackageStartupMessages(library("checkpoint"))
  suppressPackageStartupMessages(checkpoint("2021-11-30"))  # all packages in your project will be taken from that date.
}
if ("ggplot2" %in% installed.packages()){
  library("ggplot2")
} else {
  BiocManager::install("ggplot2")
  library("ggplot2")
}
if ("patchwork" %in% installed.packages()){
  library("patchwork")
} else {
  BiocManager::install("patchwork")
  library("patchwork")
}
if ("ggpmisc" %in% installed.packages()){
  library("ggpmisc")
} else {
  BiocManager::install("ggpmisc")
  library("ggpmisc")
}
if ("extrafont" %in% installed.packages()){
  library("extrafont")
} else {
  BiocManager::install("extrafont")
  library("extrafont")
}
if ("tidyverse" %in% installed.packages()){
  library("tidyverse")
} else {
  BiocManager::install("tidyverse")
  library("tidyverse")
}
if ("dplyr" %in% installed.packages()){
  library("dplyr")
} else {
  BiocManager::install("dplyr")
  library("dplyr")
}
if ("plotrix" %in% installed.packages()){
  library("plotrix")
} else {
  BiocManager::install("plotrix")
  library("plotrix")
}
if ("stringr" %in% installed.packages()){
  library("stringr")
} else {
  BiocManager::install("stringr")
  library("stringr")
}
if ("reshape2" %in% installed.packages()){
  library("reshape2")
} else {
  BiocManager::install("reshape2")
  library("reshape2")
}
```

# Get hatching per fraction

Load required functions

```{r loadfunctions include = F}
source("scripts/hatch_raw.R")
source("scripts/hatch_calc.R")
source("scripts/hatch_summ.R")
```

Plot the hatching per fraction per genotype:

```{r plothatchingperfr, fig.height = 8, fig.width = 5}
hatch_r <- hatch_raw(input = "input/210810_frac.csv")
hatch_c <- hatch_calc(input = "input/210810_frac.csv")
hatch_s <- hatch_summ(input = "input/210810_frac.csv")
hatch_s_A <- hatch_s[grep("A", hatch_s$treatment), ]
hatch_s_D <- hatch_s[grep("D", hatch_s$treatment), ]
hatch_s_S <- hatch_s[grep("S", hatch_s$treatment), ]
hatch_c_A <- hatch_c[grep("A", hatch_c$treatment), ]
hatch_c_D <- hatch_c[grep("D", hatch_c$treatment), ]
hatch_c_S <- hatch_c[grep("S", hatch_c$treatment), ]
hatch_s_W <- hatch_s[grep("water", hatch_s$treatment),]
hatch_c_W <- hatch_c[grep("water", hatch_c$treatment),]

hatch_s_A <- rbind(hatch_s_A, hatch_s_W)
hatch_c_A <- rbind(hatch_c_A, hatch_c_W)
hatch_s_D <- rbind(hatch_s_D, hatch_s_W)
hatch_c_D <- rbind(hatch_c_D, hatch_c_W)
hatch_s_S <- rbind(hatch_s_S, hatch_s_W)
hatch_c_S <- rbind(hatch_c_S, hatch_c_W)

g_A <- ggplot(hatch_s_A, aes(x = treatment, y = mean_hatch)) +
  geom_bar(stat = "identity", fill = "white", colour = "black") +
  theme_minimal() +
  geom_errorbar(aes(ymin=mean_hatch-sterr, ymax=mean_hatch+sterr), width=.2) + 
  geom_jitter(data = hatch_c_A, aes(x = treatment, y = hatchingpercent), width = 0.1) +
  scale_x_discrete(name = "") +
  scale_y_continuous(name = "% hatch", limits = c(-10,74), breaks = c(0,20,40,60,80)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1.5, hjust=1.25))
g_D <- ggplot(hatch_s_D, aes(x = treatment, y = mean_hatch)) +
  geom_bar(stat = "identity", fill = "white", colour = "black") +
  theme_minimal() +
  geom_errorbar(aes(ymin=mean_hatch-sterr, ymax=mean_hatch+sterr), width=.2) + 
  geom_jitter(data = hatch_c_D, aes(x = treatment, y = hatchingpercent), width = 0.1) +
  scale_x_discrete(name = "") +
  scale_y_continuous(name = "% hatch", limits = c(-10,74), breaks = c(0,20,40,60,80)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1.5, hjust=1.25))
g_S <- ggplot(hatch_s_S, aes(x = treatment, y = mean_hatch)) +
  geom_bar(stat = "identity", fill = "white", colour = "black") +
  theme_minimal() +
  geom_errorbar(aes(ymin=mean_hatch-sterr, ymax=mean_hatch+sterr), width=.2) + 
  geom_jitter(data = hatch_c_S, aes(x = treatment, y = hatchingpercent), width = 0.1) +
  scale_x_discrete(name = "") +
  scale_y_continuous(name = "% hatch", limits = c(-10,74), breaks = c(0,20,40,60,80)) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1.5, hjust=1.25))

g_A / g_D / g_S
```

# Metabolomics of fractions

Import, filter and trim the dfs (pos and neg)

```{r}
fr_pos <- read.csv("input/fractions_pos.csv", header = TRUE, fileEncoding = "UTF-8-BOM")
fr_neg <- read.csv("input/fractions_neg.csv", header = TRUE, fileEncoding = "UTF-8-BOM")
compounds_fr_pos <- fr_pos[,1:13]
compounds_fr_neg <- fr_neg[,1:13]
fr_pos_b <- fr_pos[,colnames(fr_pos) == "blank"]
fr_neg_b <- fr_neg[,colnames(fr_neg) == "blank"]
fr_pos <- fr_pos[,-c(1:14)]
fr_neg <- fr_neg[,-c(1:14)]
fr_pos_f <- fr_pos[(!apply(fr_pos<(fr_pos_b*1.1),1,all)),]
fr_neg_f <- fr_neg[(!apply(fr_neg<(fr_neg_b*1.1),1,all)),]
```

Do transformations (centre and scale) for PCA plot. There is no vsn transformation since there too little sample to base this transformation on.Then plot pca plots.
```{r}
library(mdatools)
fr_pos_ft <- prep.autoscale(t(fr_pos_f), center = TRUE, scale = TRUE)
fr_neg_ft <- prep.autoscale(t(fr_neg_f), center = TRUE, scale = TRUE)
meanSdPlot(t(fr_pos_f))
meanSdPlot(fr_pos_ft)
meanSdPlot(t(fr_neg_f))
meanSdPlot(fr_neg_ft)

source("scripts/plotpca.R")
plotpca2(x = fr_neg_ft) + scale_size(guide = "none")
plotpca2(x = fr_pos_ft) + scale_size(guide = "none")
```

Make plots showing counts for interesting compounds from previous correlation analyses

```{r}
# make list of compounds of interest in positive and negative mode
int_p <- c("3786")
int_n <- c("1621", "1665")
fr_n_subset <- fr_neg_f[row.names(fr_neg_f) %in% int_n, ] %>% t() %>% as.data.frame() %>% rownames_to_column("sample")
fr_n_subset <- melt(fr_n_subset, value = "sample")
fr_n_subset$genotype <- rep(c(rep("Avatar", 4), rep("Desiree", 4), rep("Seresta", 4)), 2)
fr_n_subset$fraction <- rep(c("F08","F09","F10","F11"), 6)
fr_n_subset$sample <- paste0(fr_n_subset$genotype, "_", fr_n_subset$fraction)

pn <- ggplot(fr_n_subset, aes(x = fraction, y = value)) +
  geom_line(aes(group = variable, color = variable), size = 1) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x=element_blank()) +
  ylab("counts") +
  scale_colour_discrete(name = "feature", labels = c(expression(italic(m/z)~"497.18, Rt 6.70"), expression(italic(m/z)~"525.18, Rt 7.32"))) +
  ggtitle("Negative mode") +
  facet_grid(. ~ genotype)
pn

fr_p_subset <- fr_pos_f[row.names(fr_pos_f) %in% int_p, ] %>% t() %>% as.data.frame() %>% rownames_to_column("sample")
fr_p_subset <- melt(fr_p_subset, value = "sample")
fr_p_subset$genotype <- c(rep("Avatar", 4), rep("Desiree", 4), rep("Seresta", 4))
fr_p_subset$fraction <- rep(c("F08","F09","F10","F11"), 3)
fr_p_subset$sample <- paste0(fr_p_subset$genotype, "_", fr_p_subset$fraction)

pp <- ggplot(fr_p_subset, aes(x = fraction, y = value)) +
  geom_line(aes(group = variable, color = variable), size = 1) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x=element_blank()) +
  ylab("counts") +
  scale_colour_discrete(name = "feature", labels = c(expression(italic(m/z)~"499.20, Rt 6.69"))) +
  ggtitle("Positive mode") +
  facet_grid(. ~ genotype)
pp

fr_subset <- rbind(fr_n_subset, fr_p_subset)
fr_subset$mode <- c(rep("negative", 24), rep("positive", 12))
fr_subset$label <- c(rep(expression(italic(m/z)~"497.18 (-), Rt 6.70"), 12), rep(expression(italic(m/z)~"525.18 (-), Rt 7.32"), 12), rep(expression(italic(m/z)~"499.20 (+), Rt 6.69"), 12))

# add hatch data
toMatch <- c("f08", "f09", "f10", "f11")
hatch_s_subset <- rbind(hatch_s_A[grep(paste(toMatch,collapse="|"), hatch_s_A$treatment), ],
                        hatch_s_D[grep(paste(toMatch,collapse="|"), hatch_s_D$treatment), ],
                        hatch_s_S[grep(paste(toMatch,collapse="|"), hatch_s_S$treatment), ])
hatch_s_subset$genotype <- c(rep("Avatar", 4), rep("Desiree", 4), rep("Seresta", 4))
hatch_s_subset$fraction <- rep(c("F08","F09","F10","F11"), 3)

#final plot
ppn <- ggplot(fr_subset, aes(x = fraction, y = value)) +
  geom_line(aes(group = variable, color = variable), size = 1) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x=element_blank()) +
  ylab("counts") +
  scale_color_discrete(name = "feature", 
                       labels = c(expression(italic(m/z)~"497.18 (-), Rt 6.70"), 
                                  expression(italic(m/z)~"525.18 (-), Rt 7.32"), 
                                  expression(italic(m/z)~"499.20 (+), Rt 6.69"))) +
  facet_grid(. ~ genotype)+ 
  geom_bar(data = hatch_s_subset,
               aes(x = fraction, 
                   y = (mean_hatch*25)),
               stat = "identity", 
               fill = "gray",
           alpha = 0.5) +
  scale_y_continuous(sec.axis = sec_axis(trans=~./25,
                                         name="hatching (%)",
                                         breaks = c(0, 20, 40, 60, 80)))
ppn
```

Do correlation analysis on fractionation data (correlate with hatching only)

```{r, fig.height = 8, fig.width = 8}
#correlation for positive mode
fr_pos_fc <- as.data.frame(t(fr_pos_f))
source("scripts/ordermatrixforcorplot_fr.R")
cor_fr_p <- ordermatrixforcorplot_fr(x = fr_pos_fc, compounds = compounds_fr_pos)
#correlation for negative mode
fr_neg_fc <- as.data.frame(t(fr_neg_f))
cor_fr_n <- ordermatrixforcorplot_fr(x = fr_neg_fc, compounds = compounds_fr_neg)
```

```{r, eval = F}
#get pvalues
c0_fr_p <- ordermatrixforcorplot_fr2(x = fr_pos_fc)
testRes_c_fr_p <- cor.mtest(as.data.frame(c0_fr_p), conf.level = 0.95)
testRes_c_fr_p <- as.data.frame(testRes_c_fr_p$p)
c0_fr_n <- ordermatrixforcorplot_fr2(x = fr_neg_fc)
testRes_c_fr_n <- cor.mtest(as.data.frame(c0_fr_n), conf.level = 0.95)
testRes_c_fr_n <- as.data.frame(testRes_c_fr_n$p)
```

```{r}
testRes_c_fr_p_h <- read.csv("output/testRes_c_fr_p_h.csv", header = T, fileEncoding = "UTF-8-BOM") %>% column_to_rownames("X")
names(testRes_c_fr_p_h) <- row.names(testRes_c_fr_p_h)
testRes_c_fr_p_h_subset <- get_testRes_fr(x = fr_pos_fc, compounds = compounds_fr_pos, t = testRes_c_fr_p_h)
#plot
source("scripts/plotcor.R")
corplot_fr_p <- plotcor(x = cor_fr_p, t = testRes_c_fr_p_h_subset)

testRes_c_fr_n_h <- read.csv("output/testRes_c_fr_n_h.csv", header = T, fileEncoding = "UTF-8-BOM") %>% column_to_rownames("X")
names(testRes_c_fr_n_h) <- row.names(testRes_c_fr_n_h)
testRes_c_fr_n_h_subset <- get_testRes_fr(x = fr_neg_fc, compounds = compounds_fr_neg, t = testRes_c_fr_n_h)
#plot
corplot_fr_n <- plotcor(x = cor_fr_n, t = testRes_c_fr_n_h_subset)
```


Get list of correlations
```{r}
source("scripts/correlationlist_fr.R")
c_fr_pos <- correlationlist_fr(x = fr_pos_fc, compounds = compounds_fr_pos)
c_fr_neg <- correlationlist_fr(x = fr_neg_fc, compounds = compounds_fr_neg)
```

Check if some of the high correlators from Pearson and RF are found back in any of the fractions

```{r}
compounds_fr_pos$mz_r <- round(compounds_fr_pos$m.z.meas., digits = 2)
overlap_fr_RFcorr_ph <- compounds_fr_pos[compounds_fr_pos$mz_r %in% overlap_exuph_withcorrn$mz_r, ]
compounds_fr_neg$mz_r <- round(compounds_fr_neg$m.z.meas., digits = 2)
overlap_fr_RFcorr_nh <- compounds_fr_neg[compounds_fr_neg$mz_r %in% overlap_exunh_withcorrp$mz_r, ]
overlap_fr_RFcorr_ps <- compounds_fr_pos[compounds_fr_pos$mz_r %in% overlap_exups_withcorrn$mz_r, ]
overlap_fr_RFcorr_ns <- compounds_fr_neg[compounds_fr_neg$mz_r %in% overlap_exuns_withcorrp$mz_r, ]
```
